Corrections to Manuscript No. 15-TIE-3480

### 2016-02-24 RMS/corrections.tex
create the correction file to answer the answers from reviewers in the major revision of TIE

\usepackage{alltt}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}
\newcommand{\hide}[1]{\ignorespaces}
\newcommand{\jx}[1]{{\bf Jiaxiang: }#1{ \bf End}}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}

@@

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Formal Modeling and Verification of a Rate-Monotonic Scheduling Implementation with Real-Time Maude}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc 
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

@@

% The paper headers
\markboth{IEEE Transaction on xxx,~Vol.~$x$, No.~$x$, xx~$x$}%
{Liu \MakeLowercase{\textit{et al.}}: Formal Modeling and Verification of a Rate-Monotonic Scheduling Implementation with Real-Time Maude}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

@@

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
Rate-Monotonic Scheduling (RMS) is one of the most important real-time
scheduling used in industry. There are a large number of results about
RMS, especially on its schedulability. However, the theoretical
results do not contain enough details to be used directly for an
industrial RMS implementation. On the other hand, the correctness of
such an implementation is of the crucial importance. In this paper, we
analyze a realistic RMS implementation by using Real-Time Maude, a
formal modeling language and analysis tool based on rewriting
logic. Overhead and some details of the hardware are taken into
account in the model. We validate the schedulability and the
correctness of the implementation within key scenarios. The soundness
and the completeness of our approach are substantiated.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
  Real-time systems, scheduling, embedded systems, modeling, formal
  verification, rewriting logic.
\end{IEEEkeywords}
}

% make the title area
\maketitle

% To allow for easy dual compilation without having to reenter the
% abstract/keywords data, the \IEEEtitleabstractindextext text will
% not be used in maketitle, but will appear (i.e., to be "transported")
% here as \IEEEdisplaynontitleabstractindextext when the compsoc 
% or transmag modes are not selected <OR> if conference mode is selected 
% - because all conference papers position the abstract like regular
% papers do.
\IEEEdisplaynontitleabstractindextext
% \IEEEdisplaynontitleabstractindextext has no effect when using
% compsoc or transmag under a non-conference mode.

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

%\IEEEraisesectionheading{\section{Introduction}\label{s:introduction}}
\section{Introduction}\label{s:introduction}
% Computer Society journal (but not conference!) papers do something unusual
% with the very first section heading (almost always called "Introduction").
% They place it ABOVE the main text! IEEEtran.cls does not automatically do
% this for you, but you can achieve this effect with the provided
% \IEEEraisesectionheading{} command. Note the need to keep any \label that
% is to refer to the section immediately after \section in the above as
% \IEEEraisesectionheading puts \section within a raised box.

% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps (small caps for compsoc).
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{P}{eriodic} task scheduling is one of the most important
topics within the field of industrial real-time systems.
%due to the large number of control systems that require cyclic activities~\cite{buttazzo2011hard}. 
A set of periodic tasks is said to be \emph{schedulable} with respect
to some scheduling algorithm if all jobs meet their
deadlines. \emph{Rate-Monotonic Scheduling} (\emph{RMS}) is a
\emph{fixed} priority scheduling algorithm for preemptive hard
real-time environments proposed by Liu and
Layland~\cite{DBLP:journals/jacm/LiuL73}, which assigns priorities to
jobs according to the periods of the corresponding tasks: the smaller
period, the higher priority. RMS is proven to be the \emph{optimal}
fixed priority scheduling algorithm~\cite{DBLP:journals/jacm/LiuL73},
in the sense that any set of tasks, which is schedulable under
\emph{some} fixed priority scheduling algorithm, is also schedulable
with respect to RMS. It is widely used in safety-critical real-time
applications, such as vehicles and avionics, thanks to its optimality
and easiness to implement.

Liu and Layland~\cite{DBLP:journals/jacm/LiuL73} gave a sufficient
condition for the schedulability of a set of $n$ tasks scheduled by
RMS: $\displaystyle\Sigma^n_{i=1}C_i/T_i \le n(2^{1/n}-1)$, where
$C_i$ and $T_i$ are the \emph{computation (time) requirement} and the
period of task $\tau_i$, respectively. Two main directions on RMS have
been explored since then. One is to relax the assumptions on the
original RMS model, making it applicable on more systems.  For
instance,
\cite{DBLP:conf/rtss/LehoczkySS87,DBLP:journals/rts/SpruntSL89,DBLP:conf/rtss/LehoczkyR92,DBLP:journals/tc/StrosniderLS95}
allow aperiodic tasks in the scheduling,
\cite{DBLP:journals/pe/LeungW82,audsley1993deadline} generalize RMS to
be \emph{deadline-monotonic}, \cite{DBLP:journals/tc/ShaRL90} allows
resource sharing among tasks,
\cite{dhall1978real,DBLP:journals/rts/LopezGDG03,DBLP:journals/tpds/LopezDG04,DBLP:journals/tc/BaruahG03}
extend RMS on multiprocessors, and
\cite{DBLP:journals/rts/OhS94,DBLP:journals/rts/GhoshMMS98,DBLP:journals/tpds/BertossiMR99}
enhance fault-tolerance. The other direction is to generate better
schedulability test conditions for the algorithm and its
extensions~\cite{DBLP:conf/rtss/LehoczkySD89,DBLP:conf/rtss/KuoM91,DBLP:journals/tc/BiniBB03,DBLP:journals/rts/LopezGDG03,DBLP:journals/tc/BaruahG03,gardner1999}. The
RMS algorithm is no doubt of practical importance.

It is even more crucial to ensure the reliability of an implementation
instead of the algorithm, when RMS serves in a safety-critical
system. When analyzing a realistic implementation, theoretical results
may be no more applicable. For instance, even though the conditions
derived from algorithm analysis are satisfied, schedulability can be
broken by overhead in the system, or by the interrupt mask mechanism
which may delay interrupt handling. On the other hand, correctness of
the implementation with respect to the algorithm is difficult to be
verified by the traditional methods such as testing and simulation due
to their incompleteness. Extensive effort to apply formal methods,
such as model checking and theorem proving, has been made to analyze
safety-critical systems for the past few
years~\cite{DBLP:journals/tie/MiyawakiMSYV05,DBLP:journals/iandc/MeseguerR13,DBLP:journals/cacm/Leroy09,DBLP:conf/sosp/KleinEHACDEEKNSTW09}. However,
as far as we know, few~\cite{DBLP:conf/iceccs/CuiDT14,TianD2011}
attempt to analyze the RMS algorithm, while no work for
implementations of RMS is found.

In this paper, we use \emph{Real-Time Maude}, a \emph{rewriting}-based
formal modeling language and analysis tool for real-time systems, to
model a realistic implementation of RMS, which serves as a simplified
operating system within an avionic control system, and then verify
several desired properties. Based on a realistic implementation, our
model extends the standard RMS model proposed
in~\cite{DBLP:journals/jacm/LiuL73}, by considering overhead and other
details of the hardware platform.

The rest of this paper is organized as
follows. Section~\ref{s:background} gives some background of both the
RMS algorithm and Real-Time Maude.  Section~\ref{s:imp} presents the
RMS implementation that we model and analyze.
Section~\ref{s:formalism} introduces how we model the RMS
implementation using Real-Time Maude. Then
Section~\ref{s:verification} explains how to verify the desired
properties and to evaluate the results. Related work is discussed in
Section~\ref{s:relate}. We conclude the paper in
Section~\ref{s:conclusion}.

\section{Background}
\label{s:background}

\subsection{Rate-Monotonic Scheduling Algorithm}
\label{ss:rms}
%\subsection{Formulation of the Standard Setting}
A task set consists of \emph{only} $n$ periodic tasks
$\tau_1,\ldots,\tau_n$. Each task $\tau_i$ has a period $T_i$ and a
computation requirement $C_i$. First jobs of all tasks are assumed to
be initiated at time $0$ simultaneously.  Deadlines consist of
runnability constraints only: the deadline of a job corresponding to
$\tau_i$ is the initiation time of the next job corresponding to
$\tau_i$.  The RMS algorithm chooses the labeling such that $T_1\le
T_2\le \ldots \le T_n$. Consequently, $\tau_i$ receives priority $i$,
assuming smaller numbers have higher priorities. The following
assumptions are made:

(A1) Jobs corresponding to task $\tau_i$ are initiated exactly at
times $kT_i$ with integers $k\ge 0$.

(A2) Computation requirement $C_i$ for each task $\tau_i$ is constant
and does not vary with time.

(A3) Tasks are independent, such that they are ready to run at their
initiation times and can be preempted instantly (ignoring all
blocking).

(A4) All overhead, such as task switching time, is ignored.

A simple example showing this algorithm is depicted in
Figure~\ref{sf:rmsalg}. However in this paper, we consider an
implementation instead of the RMS algorithm itself, thus the model
would be more complicated than this standard, ideal
setting. Assumption~(A1) will be modified because of the interrupt
mask mechanism, while (A4) is relaxed to obtain a more realistic
analysis model.

\subsection{Real-Time Maude}
Real-Time Maude~\cite{DBLP:journals/lisp/OlveczkyM07} is an extension
of \emph{Maude}~\cite{DBLP:journals/tcs/ClavelDELMMQ02} which is a
language and tool based on \emph{rewriting
  logic}~\cite{DBLP:journals/jlp/Meseguer12}.  It supports formal
specification and analysis of real-time systems.

\subsubsection{Specification}
Real-Time Maude models systems as \emph{modules}. A module specifies a
\emph{real-time rewrite theory} ${\cal R} = (\Sigma, E\cup A , IR,
TR)$, where:
\begin{itemize}
\item $\Sigma$ is an algebraic \emph{signature}, that is, a set of
  declarations of \emph{sorts}, \emph{subsorts} and \emph{function
    symbols}. The function symbols are allowed to be mixfix.
\item $(\Sigma, E\cup A)$ is a \emph{membership equational logic
  theory}, with $E$ a set of conditional \emph{equations} and
  \emph{memberships} on $\Sigma$, and $A$ a set of equational axioms
  such as associativity, commutativity and identity.  $(\Sigma, E\cup
  A)$ models the system's ``static'' states as an algebraic data type,
  and is equipped with a built-in specification of a sort \verb|Time|.
\item $IR$ is a set of \emph{labeled conditional rewrite rules}
  specifying the system's local transitions. Each rule has the form
  $[l]~:~s\rightarrow s'\mbox{ \textbf{if} }\bigwedge^n_{j=1}cond_j$,
  where each $cond_j$ is an equality $u_j=v_j$ and $l$ is a
  \emph{label}. Such a rule specifies an \emph{instantaneous
    transition}, without consuming time, from an instance of $s$ to
  the corresponding instance of $s'$, \emph{provided} the conditions
  hold.
\item $TR$ is a set of (\emph{labeled}) \emph{tick rules} of the form
  $[l]~:~\{s\}\rightarrow\{s'\} \mbox{ \textbf{in time} }t\mbox{
  \textbf{if} }cond$ that specify \emph{timed transitions}, which
  advances time in the \emph{entire} state $s$ by $t$ time units. $IR$
  and $TR$ together model the ``dynamic'' behaviors of the system.
\end{itemize}

In rewriting logic, rewrite rules are applied non-deterministically,
that is, when several rules can be applied on a given $s$, any of them
may be chosen. Hence non-deterministic behaviors can be modeled
naturally in Real-Time Maude.  Real-Time Maude also supports
specifications in \emph{object-oriented} style.  A class declaration
$\texttt{class }C\texttt{ |
}att_1\texttt{:}s_1\texttt{,}\ldots\texttt{,}att_n\texttt{:}s_n$
defines a class $C$ with attributes $att_1$ to $att_n$ of sorts $s_1$
to $s_n$, respectively. An \emph{object} of class $C$ is represented
as a term $\texttt{< } O\texttt{:} C \texttt{ | }
att_1\texttt{:}val_1\texttt{,} \ldots
\texttt{,}att_n\texttt{:}val_n\texttt{ >}$ of sort \verb|Object|,
where $O$, of sort \verb|Oid|, is the object's \emph{identifier}, and
$val_i$ is the value of the attribute $att_i$ with $i\in [1,n]$. A
\emph{subclass} inherits all the attributes and rules of its
superclasses.

\subsubsection{Formal Analysis}
Real-Time Maude provides many useful commands and tools to analyze a
given model. For example, \verb|rewrite| allows to execute the model,
symbolically; given an initial state, \verb|search| is used to search
reachable states satisfying the desired properties; the Maude's
\emph{Inductive Theorem Prover} (ITP) can be applied to interactively
prove properties written in \emph{membership equational logic}.

In this paper, we only consider Real-Time Maude's \emph{Linear
  Temporal Logic (LTL) model checker}, which analyzes whether
\emph{each} behavior satisfies a temporal logic formula. \emph{State
  propositions} are defined as terms of sort \verb|Prop|. Their
semantics is defined by conditional equations of the form $\texttt{ceq
} statePattern \texttt{ |= } prop \texttt{ = } b \texttt{ if } cond$,
with $b$ a term of sort \verb|Bool|, stating that $prop$ evaluates to
$b$ in states which are instances of $statePattern$ provided the
condition $cond$ holds. These equations together define $prop$ to hold
in all states $s$ that make $s \texttt{ |= } prop$ evaluate to
\verb|true|. A temporal logic \emph{formula} is constructed by state
propositions and temporal logic operators such as \verb|True|,
\verb|False|, \verb|~|(negation), \verb|/\|, \verb|\/|,
\verb|->|(implication), \verb|[]|(``always'') and
\verb|U|(``until''). Real-Time Maude supports both \emph{timed} and
\emph{untimed LTL model checking}. The untimed model checking command
\begin{alltt}
  (mc \(s\) |=u \(\Phi\) .)
\end{alltt}
checks whether the temporal logic formula $\Phi$ holds in all
behaviors starting from the initial state $s$, \emph{with no time
  limit}.

\section{The Implementation of RMS}
\label{s:imp}
The implementation is in an industrial avionic control system.
Interrupts would be triggered by, and only by, the clock every $T$,
which we call \emph{interrupt cycle}. When an interrupt request
occurs, if the system is interruptible, i.e. the interrupt mask bit is
cleared, the handler function $schedule()$ will be invoked; otherwise
$schedule()$ will be pending until the interrupt mask bit becomes
cleared.  The implementation is shown as $schedule()$ in
Figure~\ref{a:schedule}, where $taskList$ is the list of periodic
tasks to be scheduled. We assume that the list is in descending order
of priority, and both variables $taskList$ and $timer$ are global. In
this implementation, there is only one kind of interrupt, the period
$T_i$ of each task is a multiple of $T$, and the tasks are
independent, meeting assumption~(A3).

\begin{figure}[!t]
\hrule\vspace{1mm}
  \begin{algorithmic}[1]\small
\Function{$schedule$}{$ $}{}
  \State \Call{$int\_o\!f\!\!f$}{$ $}; \Comment{to disable interrupts} \label{l:1stline}
  \State \Call{$updateStatus$}{$taskList$}; \label{l:updatestatus}
  \State $timer = timer + 1$; \label{l:timer} \label{l:inc}
  \State $p = taskList$;
  \While{$p$} \label{l:startrun1st}
    \If{$p\rightarrow status == \textit{INTERRUPT}$}
      \State \Return; \label{l:return}
    \ElsIf{$p\rightarrow status == \textit{READY}$}      
      \State $p\rightarrow status = \textit{RUNNING}$;
      \State \Call{$int\_on$}{$ $}; \Comment{to enable interrupts} \label{l:endrun1st}
      \State $p\rightarrow function()$; \Comment{to execute the task} \label{l:function}
      \State \Call{$int\_o\!f\!\!f$}{$ $};
      \State $p\rightarrow status = \textit{DORMANT}$;
    \EndIf
    \State $p = p\rightarrow next$;
  \EndWhile
\EndFunction
\Function{$updateStatus$}{$p$}
  \While{$p$}
    \If{$p\rightarrow status == \textit{RUNNING}$} \label{l:startupdate}
      \State $p\rightarrow status = \textit{INTERRUPT}$;
    \EndIf
    \If{$timer~\%~(p\rightarrow period) == 0$} \Comment{the task should be initiated}
      \If{$p\rightarrow status == \textit{DORMANT}$} \Comment{the previous job finishes} 
        \State $p\rightarrow status = \textit{READY}$;
      \Else \Comment{the status is \textit{READY} or \textit{INTERRUPT}}
  \State \Call{$reportTaskError$}{$p$}; \Comment{task misses its deadline}
      \EndIf
    \EndIf \label{l:endupdate}
    \State $p = p\rightarrow next$;
  \EndWhile
\EndFunction
  \end{algorithmic}
\hrule
  \caption{The C-Like Pseudo-code of $schedule()$}
  \label{a:schedule}
\end{figure}

In Figure~\ref{a:schedule}, the handler function $schedule()$ first
updates status of all tasks in $taskList$ via function
$updateStatus()$. This updating actually initiates tasks that should
be scheduled in the current interrupt cycle. Then $schedule()$
traverses the list to execute the ready tasks one by one, or to do a
return when encountering an interrupted\footnote{Note that the status
  \textit{INTERRUPT} indicates the task is interrupted for the moment,
  or was interrupted before but its execution is not complete yet.}
task. As for the function $updateStatus()$, it updates each task in
two steps: firstly, if the task is running, it becomes interrupted;
secondly for the task at its initiation time, if its previous job is
complete, it would be set ready, otherwise it misses its deadline,
producing an error. Notice that $schedule()$ is invoked only when the
interrupt request is handled, not when the interrupt is disabled (the
interrupt mask bit is set). Due to the interrupt mask bit, its
execution cannot be interrupted when it is updating status of tasks or
searching the next task to execute, however, it can be interrupted while
executing some task (Line~\ref{l:function}). This allows the execution
of $schedule()$ to be nested.

For simplicity, in the rest of this paper, we use \emph{scheduling} to
refer to the stage, from the moment when a pending interrupt request
is detected, to the moment when the first should-be-run periodic task
starts executing, i.e. Line~\ref{l:return} or~\ref{l:function} in
Figure~\ref{a:schedule}. Therefore, \emph{scheduling time} mainly
consists of three parts: (i) the time for switching context from the
running task, possibly none, to $schedule()$ when an interrupt request
is handled, (ii) the time spent by $schedule()$ searching and setting
the \emph{first} should-be-run periodic task
(Lines~\ref{l:1stline}-\ref{l:endrun1st} in Figure~\ref{a:schedule}),
and (iii) the time for switching context from $schedule()$ to that
task. \emph{Switching} refers to the stage, from the moment when a
periodic task completes its execution, to the moment when the next
should-be-run periodic task starts executing. \emph{Switching time}
thus also consists of three parts: (i) the time for switching context
from the complete task back to $schedule()$, (ii) the time spent by
$schedule()$ searching and setting the \emph{next} should-be-run
periodic task, and (iii) the time for switching context from
$schedule()$ to that task.

\section{Formal Modeling of the Implementation}
\label{s:formalism}
Considering some technical details of the platform, such as interrupt
mask mechanism, we model the implementation under the following
assumptions:

(A1') Jobs corresponding to task $\tau_i$ are initiated at the
beginning of scheduling that handles the requests triggered at times
$kT_i$ with integers $k\ge 0$.

(A2) Computation requirement $C_i$ for each task $\tau_i$ is constant
and does not vary with time.

(A3) Tasks are independent, such that they are ready to run at their
initiation times and can be preempted instantly.

(A4') Scheduling time and switching time are considered, while other
overhead is ignored.

These assumptions make our model different from the standard one. For
instance, with assumption~(A1'), an interrupt request that occurs
during switching will be pending, such that jobs of task $\tau_i$
cannot initiate at $kT_i$. They should wait until switching finishes
and the interrupt mask bit is cleared, which is different
from~(A1). Note that (A3) says that jobs are ready to run at their
initiation times, however, no one can start running at exactly its
initiation time, because scheduling takes time. Under these
assumptions, an example showing the execution of tasks scheduled by
the implementation is depicted in Figure~\ref{sf:rmsimp}. Note that
the second job instance of $\tau_1$ is initiated at time $11$ instead
of $10$, because switching is performed and the interrupt mask bit is
set at time $10$.

\begin{figure}[!t]
\centering
\subfloat[RMS Algorithm]{
  \begin{tikzpicture}[font=\small,scale=1.25,
    % styles
    system task/.style={fill=gray!70,draw=gray},
    periodic task/.style={fill=gray!70!black,draw=gray!50!black},
    task line/.style={gray!30,very thin},
    time line/.style={gray!50,very thin,dashed},
    short line/.style={gray!50,very thin,dashed},
    init/.style={{Triangle[fill=white,scale=1.1]}-},
    complete/.style={{Triangle[scale=1.1]}-}]

  % legend
  \begin{scope}[xshift=-2cm]  
    \draw [fill=gray!25,draw=gray!50,dashed] (1.5,2.3) rectangle +(4,-1);
    \draw (1.5,2.3) node [below right] {LEGEND};
    \draw [init] (2,1.85) -- +(0,-0.01);
    \draw (2.1,2) node [below right] {: Initiation of Tasks};
    \draw [complete] (2,1.42) -- +(0,0.1);
    \draw (2.1,1.7) node [below right] {: Completion of Tasks};
  \end{scope}

    % grid
    \foreach \x in {0.3,0.7}
      \draw[task line] (0,\x) -- (5.2,\x);

    % axes
    \draw[->,thick] (0,0) node [below] {$0$} -- (5.3,0) node [above] {time};
    \draw[thin] (0,0) -- (0,1);

>>>>

\usepackage{alltt}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}
\usepackage{multicol}
\newcommand{\hide}[1]{\ignorespaces}
\newcommand{\jx}[1]{{\bf Jiaxiang: }#1{ \bf End}}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}

@@

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Corrections to Manuscript No. 15-TIE-3480}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc 
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

@@

% The paper headers
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

@@

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.

% make the title area
%\maketitle

% To allow for easy dual compilation without having to reenter the
% abstract/keywords data, the \IEEEtitleabstractindextext text will
% not be used in maketitle, but will appear (i.e., to be "transported")
% here as \IEEEdisplaynontitleabstractindextext when the compsoc 
% or transmag modes are not selected <OR> if conference mode is selected 
% - because all conference papers position the abstract like regular
% papers do.
%\IEEEdisplaynontitleabstractindextext
% \IEEEdisplaynontitleabstractindextext has no effect when using
% compsoc or transmag under a non-conference mode.

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
%\IEEEpeerreviewmaketitle

\twocolumn[
  \begin{@twocolumnfalse}
\maketitle

\begin{center}
  \begin{tabular}{r l}
    \hline\hline
    & \\
    Manuscript Number: & 15-TIE-3480\\
    Manuscript Title: & Formal Modeling and Verification of a Rate-Monotonic Scheduling Implementation\\
    & with Real-Time Maude \\
    Submitted to: & Transactions on Industrial Eletronics \\
    Manuscript Type: & Regular paper \\
    & \\
    \hline\hline
  \end{tabular}
  \bigskip
\end{center}

  \end{@twocolumnfalse}
]

\section{Answers to the Reviewers}
\subsection{Reviewer 1}
Comments to the Author

The paper shows a realistic implementation of a Rate-Monotonic
Scheduling algorithm using Real-Time Maude (a modeling language based
on rewriting logic) by taking into account the overhead of scheduling
and some other details of the hardware platform. The correctness of
the implementation with respect to the algorithm and the completeness
of the model are verified by model checking, and validated within
different realistic scenarios.

On the positive side, this paper presents a solid piece of
work. Periodic task scheduling is an important (and difficult)
problem, and the topic of the paper appears to be well suited within
the field of the industrial real-time systems. Overall the paper is
well structured, technically accurate, presents the work in a
comprehensive and reasonable way, and tackles theoretical as well as
practical aspects. There are two principal contributions of the paper:

1) A novel implementation of the Rate-Monotonic Scheduling algorithm,
which contains more complex and realistic details instead of the ideal
setting.

2) Rate-Monotonic Scheduling is investigated using Real-Time Maude for
the first time, to apply formal methods such as model checking and
theorem proving to analyze theoretical results, verify desired
properties, and evaluate results.

On the negative side:

1) The weak point is the lack of evidence of practicality, efficiency
and scalability. Since there are only small and simplistic scenarios
to analyze this model, it is difficult to judge the effectiveness of
the proposed implementation. The key scenarios appear conveniently
small to allow the implementation fully supports all these examples,
returning at the same time appropriate counterexamples to guide and
adjust the design. Is this true for larger examples of "real world"
applications within the field of industrial real-time systems? How do
the authors propose to handle very large scenarios? ¿How does it
behave on typical real-time scheduling problems used in industry? I
would like to see more complex examples to convince the reader of the
practicability of the approach on the assumptions given for the model,
and would probably increase the mentioned benefits. The paper does not
give any directions to the above problems, and some sentences pointing
out the difficulties in the implementation of more generic and
realistic scenarios would be most welcome to understand the real
contribution of the paper.

2) One of my biggest problems with the theoretical part of this work
is that the paper is technically solid as far as I can tell, but
there's not much yet in the way of theorems and results about formal
properties in Maude. The complete and detailed proof of the Theorem 2
should be provided somewhere. If the proof of the main results of
schedulability and correctness of the model is straightforward (as is
suggested) it seems not necessary to present a complex formalism on
Real-Time Maude and all the theorems provided are natural consequences
of Theorem 1. I agree that even non-surprising results need to be
investigated in the proof of Theorem 2, some sentences pointing out
the difficulties in the theoretical formalization in Real-Time Maude
would be most welcome. Please, include (e.g., in an appendix or even
in a technical report) a detailed proof.

3) Another important point which needs improvement is the description
of the implementation. I guess other readers might have less
difficulties with the motivations if they know Maude and the Real-Time
Maude extension, but if you aim for a more general audience, it might
not work. Many pages are spent for the description of the formal
modeling of the implementation and the section is hard to read. I
highly recommend the authors to thoroughly revise this part of the
paper to make it easy-to-read. It would be better preferable to have a
much shorter description of the principal insights for the
implementation followed by more convincing examples and benchmarks
with respect to related work to convince the reader of the
practicability of the approach. You should try to summarize this
section and to extend the original contribution. From my point of
view, there is a trade off in this paper between the contributions and
the presentation of the implementation.

In conclusion, the RMS implementation proposed by the authors by using
Real-Time Maude is interesting, but not complete (theorems and results
about the formal modeling and verification of the implementation in
Maude ought be clearly demonstrated). Moreover, the weak point is the
lack of evidence for practicality, and the presentation of Section IV
can be improved. It's not clear at all how this kind of key scenarios
could actually run in practice in more general real-time
systems. However, I think that the paper is worthy of publication,
because is the first time that Rate-Monotonic Scheduling is
investigated using Real-Time Maude in a (more) realistic
implementation, providing a solid semantic foundation for formal
verification. Therefore, I recommend the (weak) acceptance of this
paper.

Some minor remarks:

- Page 7, Related Work. More comparisons with other approaches based
on model checking would help to evaluate pros and cons of an
implementation with Real-Time Maude with respect to different
languages and tools. The authors should review the comparison with
[24] and [25] more thoroughly. You should benchmark your approach
against others.

- Page 8, Conclusions. Where in the paper do you show that the details
of your "realistic" implementation are "sufficient" for the behaviors
of all real systems used in the current industry? This should be
briefly discussed.

\subsection{Reviewer 2}
Comments to the Author

The paper proposes a new way of modelling Rate Monotonic Scheduling
(RMS).  The new modelling approach is more accurate than the
simplified way originally proposed by Liu and Layland, as it also
models some of the overhead that is present in a real system
(e.g. scheduling and switching time).  This implies taking some of the
hardware details into account.  The model, which is based on the
formal language of Real-Time Maude, can be analyzed using the formal
verification tools provided by Real-Time Maude.  The paper shows how
this approach can be used for modelling and analyzing a particular RMS
implementation.

Up to my knowledge, this is original work. Related papers address
different aspects and do not take implementation overhead into
account.

The technique proposed by the paper is of industrial interest because
it improves the accuracy of analyzing schedulability in RMS
implementations with respect to the bare application of the theoretic
schedulability result.

The paper is well written and it looks sound, but I could not check it
completely.

One of the problems with this paper is that it is not completely
self-contained.  Some knowledge about Real-Time Maude is necessary in
order to fully understand it.  I found the short introduction about
Real-Time Maude (section II.B) not enough for understanding the
formalism used in the rest of the paper (see details below).

Another concern is that the significance of the paper could be much
better if the paper addressed the problem of modelling RMS in a more
general way. Instead, the paper takes the form of a case study, where
the formal modelling of one specific RMS implementation is presented,
rather than a general method for modelling and analyzing RMS
implementations.  Also, the discussion about soundness and
completeness of the analysis (section V.C) is not fully developed and,
once again, it refers to the specific model rather than being a
general result.

As a conclusion, even if the significance of the paper is not as much
as it could be, I think the paper provides something novel and
interesting for the readers. Hence, my recommendation is to accept
this paper provided it is made more self-contained (without assuming
any reader's knowledge about Maude) and the proof of completeness and
soundness is fully developed.

Detailed comments

Section II.B \\ In the definition of IR, it is not clear what s and s'
are. I guess they are states, but the concept of state was not
introduced. \\ In the definition of a class, the authors should
specify that a class includes a collection of rules, otherwise this
fact may be missed.

Section III \\ Page 2 right column, line -24 The implementation is
shown as schedule() in Figure 1 -> The pseudocode of schedule() is
shown in Figure 1

Section IV.A \\ I don't understand the meaning of [ctor] in the
definition of some operations (maybe because I have no specific
background on Maude). I could not find the explanation of this writing
in the introduction to Maude. \\ For the stack sort, the authors
mention operations push, pop and peek, but then you don't write the
definitions of these operations. Instead, they write the definitions
of bottom and \#, which are not explained in the text. This part should
be made clearer. \\ Also, the meaning of NzNat is not explained.

Section V.C \\ In the statement of theorem 1, the definition of
time-robust real-time rewrite theory is missing. Also, the meaning of
tick-stabilizing atomic propositions is not defined. \\ This makes the
section not self-contained. \\ The proof of theorem 2 is missing. The
authors just indicate how the proof could be developed, but they do
not develop it. A possibility would be to point to a document where
this proof has been developed.

\subsection{Reviewer 3}
Comments to the Author

This paper stuides the implementation of a rate-monotonic scheduling
with real-time maude, according to my carefully reading of this paper,
I do not think this paper is qualified for TIE due to the following
reasons:

1. Rate-monotonic scheduling is a very simple scheduling and has been
sufficiently invesigated in the field of embedded system, the authors
claim the innovation of this paper as the implementation using
real-time maude, however, I cannot see any special point (or
advantage) when using this real-time maude in modelling RMS.

2. RMS has many different realizatio methods, no comparison is given
to justify the strong point of real-time maude when applied to RMS.

3. The assumptions given in IV is too rigid, which makes the
scheduling problem to be investigated by very simple, then what is the
difficulty?

4. The description in section IV (A-E) is trivial, it can only be
viewed as a simple application of the rewriting logic.

5. The model checking part is not sufficiently discussed, the authors
only list the commands but not illustrate the mechanism behind these
commands.

Based on the above comments, I thinks this paper is only a simple
application [26] and [27] in RMS, hence I do not recommend this paper.
